---
title: "2024 Posit Table Competition Entry"
author: "Diarmuid Lloyd"
format: 
 html:
     page-layout: full
editor: visual
execute: 
  warning: false
  message: false
  echo: true
---

## ðŸ´ó §ó ¢ó ³ó £ó ´ó ¿ âš½ï¸ Introduction

Selected 2023-24 statistics of the top-flight league of professional Scottish football, the Scottish Premiership.

The table is built from three scripts run in sequence, although the first can be skipped after the first run:

1.  `0_collect_timeseries.R`: Collecting time series data back to the 2000-01 season.

2.  `01_spfl_collect_data.R`: Wrangling the data, and calculating summary statistics

3.  `02_build_table_gt.R`: Building the table

[The repo with these scripts can be found here.](https://github.com/dplloyd/posit-table-comp-2024)

```{r}

library(tidyverse)

source("0_collect_timeseries.R")

source("01_spfl_collect_data.R")

source("02_build_table_gt.R")
```

## Data source

Data are sourced from [https://www.football-data.co.uk/scotlandm.php](https://www.football-data.co.uk). While no specific licence is provided on the website, the maintainers state:

> You are free experiment with the data yourselves, but if you are looking for a bespoke Excel application that has been desinged specifically to work with Football-Data's files, visitÂ Â [BetGPS](http://www.betgps.com/betting-data.php)Â for an exceptional data analysis workbook. Like all of Football-Data's files, it free to download.

The [accompanying notes](https://www.football-data.co.uk/notes.txt) provide full sources for the data, but are summarised as:

> ```         
> Current results (full time, half time)
> XScores - http://www.xscores .com
>
> Match statistics
> BBC, ESPN Soccer, Bundesliga.de, Gazzetta.it and Football.fr
>
> Bookmakers betting odds
> Individual bookmakers
>
> Betting odds for weekend games are collected Friday afternoons, and on Tuesday afternoons for midweek games.
>
> Additional match statistics (corners, shots, bookings, referee etc.) for the 2000/01 and 2001/02 seasons for the English, Scottish and German leagues were provided by Sports.com (now under new ownership and no longer available).
> ```

## Table

The table is built in `gt` and `gtExtra`. I relied heavily on the latter's help files, and [Albert Rapp's](https://gt.albert-rapp.de) tutorial on creating plots within table cells, and I particularly wanted to explore how sparklines and other in-cell plots can help summarise tabular data. While I didn't exploit the functionality of embedding `ggplot2` objects in cells, I think that route would help in future iterations, and exploring other variables to summarise.

```{r}
final_table
```
